Reproducible and programmatic generation of neuroimaging visualisations

### 2022-03-18 manuscript.Rmd
sp check

---
title: "Reproducible and programmatic generation of neuroimaging visualisations"
author: "Sidhant Chopra & [coauthors]"
date: "09/03/2022"
output:
  pdf_document: default
  html_document: default
editor_options: 
  markdown: 
    wrap: 72
---

@@

Visualising neuroimaging data is one of the primary ways in which we
evaluate data quality, interpret results and communicate findings. These
visualisations are commonly produced using graphic interface-based (GUI)
tools where individual images are opened and display settings are
manually changed until the desired output is reached. While the choice
to use GUI-based software's has been driven by a perception of
convenience, flexibility and accessibility, there are now many
code-based software packages which are well-documented and don't often
require high-level knowledge of the programming. These tools are
flexible and allow for generation of reproducible, high-quality and
publication ready brain visualisations in only a few lines of code
(Figure 1), especially within the R, Python and Matlab environments.

@@

Often programming languages such as R, Python and Matlab are used for
the analysis in neuroimaging studuies, but the brain visualisation
resulting from these analyses is outsourced to a septate GIU-based
visualisation tools such as FSLeyes, Freeview or ITK-snap. Switching
from analysis to a GUI-based visualisation process be a cumbersome
deviation from the scientific workflow and can make debugging errors
more difficult, as you have to regularly switch program to visually
examine the results of any modifications to analyses. Using the brain
visulitsation ools that already exist within your chosen programming
environment can provide instant visual feedback on the impact of
modifications to processing or analyses. While some GUI-based tools do
offer command-line access to generate visualisations, they can lack
flexibility to easily generate publication ready figures and can lose
the benefits provided by your preferred programming environment.

@@

Neuroimaging data are often spatially 3D and can have multiple time
points, adding a 4th dimensional (e.g. fMRI data). Thus, communicating
findings or evaluating quality using using static 2D slices is
challenging. While well-curated 3D renderings can help with spatial
localisation (cite madan), in the end, static images can only provide an
incomplete representation of the data. A added advantage of some of
code-based tools is that you can generate 'rich' media like interactive
figures or animations, which can improve scientific communication of
findings. Many code-based software tools now allow for the generation of
animated brain visualisations (cite blog) or interactive widgets which
allow users to zoom, rotate and scroll through slices. Linking to or
even embedding these videos or interactive figures in papers can greatly
enhance the communication of findings.

>>>>

---
title: "Reproducible and programmatic generation of neuroimaging visualisations"
author: "Sidhant Chopra & [coauthors]"
date: "09/03/2022"
output:
  word_document: default
  pdf_document: default
  html_document: default
editor_options:
  markdown:
    wrap: 72
---

@@

Visualising neuroimaging data is one of the primary ways in which we
evaluate data quality, interpret results and communicate findings. These
visualisations are commonly produced using graphic interface-based (GUI)
tools where individual images are opened and display settings are
manually changed until the desired output is reached. While the choice
to use GUI-based software's has been driven by a perception of
convenience, flexibility and accessibility, there are now many
code-based software packages which are well-documented and don't often
require high-level knowledge of programming. These tools are
flexible and allow for generation of reproducible, high-quality and
publication ready brain visualisations in only a few lines of code
(Figure 1), especially within the R, Python and Matlab environments.

@@

Often programming languages such as R, Python and Matlab are used for
the analysis in neuroimaging studies, but the brain visualisation
resulting from these analyses is outsourced to a septate GIU-based
visualisation tools such as FSLeyes, Freeview or ITK-snap. Switching
from analysis to a GUI-based visualisation process be a cumbersome
deviation from the scientific workflow and can make debugging errors
more difficult, as you have to regularly switch program to visually
examine the results of any modifications to analyses. Using the brain
visualisation tools that already exist within your chosen programming
environment can provide instant visual feedback on the impact of
modifications to processing or analyses. While some GUI-based tools do
offer command-line access to generate visualisations, they can lack
flexibility to easily generate publication ready figures and can lose
the benefits provided by your preferred programming environment.

@@

Neuroimaging data are often spatially 3D and can have multiple time
points, adding a 4th dimensional (e.g. fMRI data). Thus, communicating
findings or evaluating quality using static 2D slices is
challenging. While well-curated 3D renderings can help with spatial
localisation (cite madan), in the end, static images can only provide an
incomplete representation of the data. A added advantage of some of
code-based tools is that you can generate 'rich' media like interactive
figures or animations, which can improve scientific communication of
findings. Many code-based software tools now allow for the generation of
animated brain visualisations (cite blog) or interactive widgets which
allow users to zoom, rotate and scroll through slices. Linking to or
even embedding these videos or interactive figures in papers can greatly
enhance the communication of findings.
